---
title: "Biostat 200C Homework 5"
subtitle: Due 11:59PM June 2nd
author: Jiahao Tian
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
library(tidyverse)
library(faraway)
library(mgcv)
```

## Q1. Doctor visits in Australia, ELMR Exercise 13.4 

```{r}
data(dvisits)
help("dvisits")
```

The `dvisits` data comes from the Australian Health Survey of 1977–78 and consist of
5190 single adults where young and old have been oversampled. Use `help("dvisits")` to check the variables.

(a) Build a generalized additive model with `doctorco` as the response and `sex`, `age`,
`agesq`, `income`, `levyplus`, `freepoor`, `freerepa`, `illness`, `actdays`, `hscore`, `chcond1` and `chcond2` as possible predictor variables. Select an appropriate size for your model. (Hint. fit a simpler model first and check some marginal plots.)

**Solution:**

```{r}
summary(dvisits)
```

```{r}
dvisits %>%
  ggplot(mapping = aes(x = actdays, y = doctorco)) +
  geom_point(size = 1) +
  geom_smooth() +
  theme_bw()

dvisits %>%
  ggplot(mapping = aes(x = freepoor, y = doctorco)) +
  geom_point(size = 1) +
  geom_smooth() +
  theme_bw()

#Lots of binary, no need to do transformation.
```

```{r}
#It is not necessary to use age squared term, when we used age term and transformed it.

dvgam <- gam(doctorco ~ sex + s(age) + s(income) + levyplus + freepoor +
               freerepa + factor(illness) + s(actdays) + s(hscore) + 
               chcond1 + chcond2, family = poisson, scale = -1, data = dvisits)
summary(dvgam)
```

The variables which are not significant are `levyplus, chcond1, chcond2, freerepa`

The final model; 
`doctorco ~ sex + age + s(income) + freepoor + factor(illness) + s(actdays) + s(hscore)`


(b) Check the diagnostics.

**Solution:**

```{r}
dvgam2 <- gam(doctorco ~ sex + age + s(income) + factor(illness) +
                s(actdays) + s(hscore) + freepoor, 
              family = poisson, scale = -1, data = dvisits)
anova(dvgam2, dvgam, test = "Chisq")
```

From the p-value 0.5365, which tells that there is not a significant differences between these two model,
both of them explain the same variances.

```{r}
dvisits %>%
mutate(devres = residuals(dvgam2, type = "deviance"),
linpred = predict(dvgam2, type = "link")) %>%
ggplot + geom_point(mapping = aes(x = linpred, y = devres)) +
  labs(x = "Linear predictor", y = "Deviance residual")
```

```{r}
plot(dvgam2, residuals = TRUE)
```


(c) What sort of person would be predicted to visit the doctor the most under your
selected model?

**Solution:**

```{r}
summary(dvgam2)
```


We see that the type of patient most likely to visit a doctor:

- female
- older people
- low income
- has recent illness
- has reduced recent days of activity due to injury or illness
- has bad health according to the Goldberg scale
- is covered free by government because of old-age or disability pension
- is not covered by government because of lower income


(d) For the last person in the dataset, compute the predicted probability distribution for
their visits to the doctor, i.e., give the probability they visit 0,1, 2, etc. times.

**Solution:**

```{r}
(lastp <- predict(dvgam2, type = "response")[nrow(dvisits)])
lp <- round(dpois(0 : 4, lastp), 3)
names(lp) <- 0 : 4
lp
```



